External priors for the next generation of CMB experiments.

### 2016-02-24 Notes/response.txt

Major comments:

- In my opinion the authors are over-discussing well known
degeneracies between cosmological parameters adding no new
information. They formulate the key question in the second-to-last
paragraph of the introduction but there is no answer at the end of the
reading. While acknowledging the importance of having priors on some
parameters to better constrain neutrinos and dark energy, I think the
authors should explain or highlight how those priors will be obtained,
i.e., which external experiments are needed and how they will help CMB
Stage-IV. Moreover, additional datasets like galaxies surveys have
already been proved to be useful by a number of papers in the
literature, the authors should highlight what is new in their results.
Following on these concerns, it is not clear to me if the authors have
an idea on how to get those improvements, for example how to get a
better prior on n_s with non-CMB data. I find the discussion simply
speculative. This impacts in particular the abstract, the results and
the conclusion sections.

The key issue raised here seems to be: How best to address the question: What external information can be used to improve projected constraints by breaking degeneracies?
We think the referee is saying there are 2 parts to this answer: (i) quantify how much the added information on other parameters breaks degeneracies and (ii) give a sense of how we could get that information. The referee points out that we did not do a good job explaining the possibilities for (ii) and that for (i) at least qualitatively other papers have discussed degeneracies. We think this paper does advance the discussion of how degeneracies impact future constraints. The plots in Figs. 1-4 are a new way of answering this question and we think quite useful. We thought it was important to explain the degeneracies in the text for completeness, but we agree with the referee that these explanations are not new. It is the quantitative information in the plots that is new.

### 2016-02-24 Notes/CMBS4_priors_paper.tex
very minor changes

	
% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showpacs' option to make PACS codes appear
%  Add 'showkeys' option to make keywords appear
\documentclass[aps,prd,reprint,superscriptaddress,nofootinbib]{revtex4-1}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{graphicx}
\usepackage{color}

%\newcommand\sd[1]{{\bf SD: #1}}
\newcommand{\sd}[1]{\textcolor{red}{[{\bf SD}: #1]}}
\newcommand{\am}[1]{\textcolor{blue}{[{\bf AM}: #1]}}
\newcommand{\yp}[1]{\textcolor{green}{[{\bf YP}: #1]}}

%\newcommand\refsec[1]{\S\ref{sec:#1}}
%\newcommand\refeq[1]{Eq.~(\ref{eqn:#1})}
\newcommand\refsec[1]{\S\ref{sec:#1}}
\newcommand\refeq[1]{Eq.~(\ref{eqn:#1})}
\newcommand{\refssec}[1]{section~\ref{subsec:#1}}
\newcommand{\reffig}[1]{Fig.~\ref{fig:#1}}
%\newcommand{\apj}{ApJ}
\newcommand{\physrep}{Physics Reports}
\newcommand{\aapr}{A\&A Rev.}
\newcommand{\apjl}{ApJL}
\newcommand{\apjs}{ApJS}
\newcommand{\aap}{A\&A}
\newcommand{\mnras}{MNRAS}
%\newcommand{\prd}{Phys. Rev. D}
\newcommand{\physrev}{Phys. Rev.}
\newcommand{\physrevlett}{Phys. Rev. Lett.}
\newcommand{\jcap}{{\em JCAP }}
\newcommand{\aj}{AJ }
\bibliographystyle{apsrev4-1}

@@

\section{Results \label{sec:results}}
In this section we present our main results. We will start from $\nu \Lambda$CDM and then explore extensions to this model by considering constraints on $N_{\rm eff}$ and then $w$.

@@

Small scale structure formation is suppressed if fast-moving neutrinos comprise a significant part of the matter budget. 
Indeed, below the free-streaming scale, the matter power spectrum is suppressed in the presence of
massive neutrinos by a factor $\Delta P/P\simeq -8f_{\nu}$ , where $f_{\nu} = \Omega_{\nu} / \Omega_{m}$ is the contribution of neutrinos to the total matter density.
This effect is probed by CMB lensing and, since the active neutrino number densities are known in the Standard Model, can be directly transformed into a constraint on the sum of the neutrino masses $\sum m_\nu$. This is particularly exciting because we now know that neutrinos are massive, with a lower limit $\sum m_\nu>50$ meV that emerges from oscillation experiments (see \cite{bellini:2013} and references therein).

%==========
\begin{figure}[htbp]
\begin{center}
\includegraphics{{prior_omnuh2_lmin=4_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
\caption{Projected constraints on the sum of the neutrino masses (absolute on the left and relative on the right) from CMB-S4 and current BAO measurements as a function of priors on the other cosmological parameters. The {\it internal} (marginalized constraints from CMB-S4 and current BAO) constraints on each parameter are given as $\sigma_{\rm pipeline}$. The horizontal (blue) line correspond to CMB S4 experiment + Planck Pol + DESI.}
\label{fig:prior_omeganuh2-h}
\end{center}
\end{figure}
%==========

%%==========
%\begin{figure}[htbp]
%\begin{center}
%\includegraphics{{prior_omnuh2_notau_lmin=50_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
%\caption{Constraints on the sum of the neutrino masses (absolute on the left and and relative to its value on the right) as a function of priors on $\Omega_{c} h^{2}$. The horizontal (blue) line correspond to CMB S4 experiment + Planck Pol + DESI.}
%\label{fig:prior_omeganuh2-omch2}
%\end{center}
%\end{figure}
%%==========

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics{{prior_omnuh2_lmin=50_lmax=4499_ndet=1000000_fsky=0.75_test}.pdf}
\caption{Same as Fig.~1, but with $l<50$ modes excluded from CMB-S4, instead coming from Planck. Note the importance of obtaining information on the optical depth $\tau$ if the low-$l$ modes are not measured accurately.}
\label{fig:prior-50}
\end{center}
\end{figure}
%==========
%
%%==========
%\begin{figure}[htbp]
%\begin{center}
%\includegraphics{{prior_omnuh2_hubble_lmin=50_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
%\caption{Constraints on the sum of the neutrino masses (absolute on the left and and relative to its value on the right) as a function of priors on the Hubble parameter $H_{0}$. The horizontal (blue) line correspond to CMB S4 experiment + Planck Pol + DESI.}
%\label{fig:prior_omeganuh2-50-h}
%\end{center}
%\end{figure}
%%==========
%
%
%%==========
%\begin{figure}[htbp]
%\begin{center}
%\includegraphics{{prior_omnuh2_omch2_lmin=50_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
%\caption{Constraints on the sum of the neutrino masses (absolute on the left and and relative to its value on the right) as a function of priors on $\Omega_{c} h^{2}$. The horizontal (blue) line correspond to CMB S4 experiment + Planck Pol + DESI.}
%\label{fig:prior_omeganuh2-50-omch2}
%\end{center}
%\end{figure}
%%==========
%
%%==========
%\begin{figure}[htbp]
%\begin{center}
%\includegraphics{{prior_omnuh2_re_optical_depth_lmin=50_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
%\caption{Constraints on the sum of the neutrino masses (absolute on the left and and relative to its value on the right) as a function of priors on the optical depth $\tau$. The horizontal (blue) line correspond to CMB S4 experiment + Planck Pol + DESI.}
%\label{fig:prior_omeganuh2-50-tau}
%\end{center}
%\end{figure}
%%==========
%

\reffig{prior_omeganuh2-h} shows the projected errors on $\sum m_\nu$ from CMB-S4 (+Planck+current BAO) as a function of external priors on $H_{0}, \Omega_{c}h^{2}, n_s$ and $A_s$. 
Without any external priors (far right in figure), upcoming CMB experiments are poised to obtain 1-sigma limits on $\sum m_\nu$ of order 25 meV, corresponding to a 2-sigma detection even in the worst case scenario. Adding in external priors helps this significantly though, as discussed, for example, in \cite{di-valentino:2015,2013arXiv1309.5383A,pan:2015a,allison:2015}. 
There, the added prior was ``DESI BAO'', representing a measurement of distances as a function of redshift from the baryon acoustic oscillation feature probed by the planned experiment DESI \cite{levi:2013}. This is represented by the blue band in our figures. In addition to showing the implications of introducing this fixed external constraint, \reffig{prior_omeganuh2-h} generalizes the extent to which external information will improve the neutrino mass constraints. For example, in \reffig{prior_omeganuh2-h} the red solid curve shows the improvement as additional information on the Hubble constant is applied. We can see, for example that, imposing {\it external} information at the level of $\sigma_{H_0}/H_0=0.5\%$ would improve the neutrino mass constraint to 17 meV. It is not surprising that this is close to the constraint obtained with DESI because the BAO distances in this model depends most sensitively on $\Omega_ch^2$ and $H_{0}$.
Similarly, adding to CMB S4 + BAO15 an external constraint on the dark matter density equal to its internal constraint of $0.4\%$ improves the projected neutrino mass constraint to 20 meV. These are multi-dimensional examples of the familiar fact that adding two sets of equal constraints (in this case internal to S4+BAO15 and external) reduces the overall error by a factor of $\sqrt{2}$.

@@

%\begin{figure}[htbp]
%\begin{center}
%\includegraphics{prior_omnuh2_50.pdf}
%\caption{Same as \reffig{prior_omeganuh2} but with $\ell^{\rm CMB-S4}_{\rm min}>50$. Here $A_{s}$ and $\tau$ priors will also helps because of their degeneracy with $\sum m_{\nu}$.}
%\label{fig:prior_omeganuh2_tau}
%\end{center}
%\end{figure}

Because of the presence of foregrounds, it may be difficult to get the very low-$\ell$ modes from the ground with CMB-S4, so in \reffig{prior-50}, we limit the CMB S4 modes to $\ell>\ell_{\rm min}=50$. For the large scale modes $\ell<\ell_{\rm min}=50$ we use Planck at the level of accuracy forecasted by the Planck Blue Book.  
Together with the importance of $H_{0}$ and $\Omega_ch^2$ previously observed, in this case external priors on $A_{s}$ and $\tau$ are found to be important. % (see \reffig{prior_omeganuh2-50-ns} and \reffig{prior_omeganuh2-50-as}).

Indeed, as recently pointed out by \cite{allison:2015}, a degeneracy exists between $\sum m_\nu$ and $A_{s}$ and, as a consequence, between $\sum m_\nu$ and $\tau$. This can be understood as follows.
The CMB lensing reconstruction is noisy at scales larger than the neutrino free streaming scale (see Fig. 5 of \cite{2013arXiv1309.5383A}). For this reason it is hard to measure the unsuppressed CMB lensing power spectra that is then compared to the small scale suppressed one to constrain the values of $f_{\nu}$. The unsuppressed amplitude of the lensing spectrum is actually better constrained by the primordial amplitude $A_{s}$, which is probed by the primordial CMB. Unfortunately, the CMB spectra are sensitive to $A_{s}\exp^{-2\tau}$, so a measurement of the optical depth is required to infer $A_{s}$ and therefore to tighten the constraints on $\sum m_\nu$. If CMB S4 will not be designed to measure scales $\ell<50$ to avoid difficulties in dealing with atmospheric emission, an external prior on the optical depth will be crucial. This can be obtained from future 21 cm surveys \cite{liu:2015} or other CMB experiments, satellite like LiteBIRD \cite{matsumura:2014}, PIXIE\cite{kogut:2011}, and COrE/PRISM \cite{the-core-collaboration:2011,prism-collaboration:2013}, balloon-borne experiments like SPIDER \cite{filippini:2010}, EBEX \cite{reichborn-kjennerud:2010} or eventually ground experiments like CLASS \cite{essinger-hileman:2014}.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics{{prior_massless_neutrinos_lmin=4_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
\caption{Constraints on $N_{\rm{eff}}$ (absolute on the left and and relative to its value on the right) as a function of priors on the $\Omega _{b} h^{2}$. The horizontal (blue) line correspond to CMB S4 experiment + Planck Pol + DESI.}
\label{fig:prior_neff-ombh2}
\end{center}
\end{figure}

%\begin{figure}[htbp]
%\begin{center}
%\includegraphics{{prior_massless_neutrinos_omch2_lmin=4_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
%\caption{Constraints on $N_{\rm{eff}}$ (absolute on the left and and relative to its value on the right) as a function of priors on $\Omega_{c} h^{2}$. The horizontal (blue) line correspond to CMB S4 experiment + Planck Pol + DESI.}
%\label{fig:prior_neff-omch2}
%\end{center}
%\end{figure}
%
%\begin{figure}[htbp]
%\begin{center}
%\includegraphics{{prior_massless_neutrinos_scalar_spectral_index(1)_lmin=4_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
%\caption{Constraints on $N_{\rm{eff}}$ (absolute on the left and and relative to its value on the right) as a function of priors on $n_{s}$. The horizontal (blue) line correspond to CMB S4 experiment + Planck Pol + DESI.}
%\label{fig:prior_neff-ns}
%\end{center}
%\end{figure}

\reffig{prior_neff-omch2}, \reffig{prior_neff-ns} and  \reffig{prior_neff-ombh2} show projections for how well CMB-S4 will do at measuring $N_{\rm eff}$ as a function of priors on $H_{0}$, $\Omega_{c}h^{2}$ and $\Omega_{b}h^{2}$(the sum of the neutrino masses must be included as a free parameter). With no priors on the other parameters, the projected $1$-sigma error is $0.016$, close to a 3-sigma detection of the expected deviation from $N_{\rm eff}=3$.

The sensitivity to relativistic degrees of freedom comes from the effect of extra species on the damping tail of the CMB anisotropies \cite{2013PhRvD..87h3008H}, both in temperature and polarization, so parameters that also strongly affect this part of the spectrum, like the slope $n_s$ and the baryon density $\Omega_bh^2$, are the most relevant to improve the $N_{\rm eff}$ constraint. As such, \reffig{prior_neff-ombh2} and \reffig{prior_neff-ns} show that obtaining external priors on either of these would reduce the errors on $N_{\rm eff}$ to get close to a 4-sigma detection of the partial decoupling prediction. Besides futuristic high redshift 21 cm ideas, the best bet to improve the constraints on $n_s$ are probes that push to redshift $z>1$ such as the Lyman alpha forest. There are many more linear Fourier modes available at high $z$ that can be used to obtain the larger level arm from large to small scales that is essential to extract high precision measurements of $n_s$.

@@

\reffig{prior_w-h} shows that without any priors, CMB experiments (+BAO15) will do better than current constraints, which hover around 10\%. However, an external constraint on the Hubble constant would improve the CMB constraints on $w$ by a factor of 2. 
This is primarily due to the fact that CMB constrains $w$ through a precise measurement of the comoving distance to the last scattering surface. However that can be kept fixed while varying $w$ by accordingly changing the value of $H_{0}$ (and $\Omega_{\Lambda}$ to keep the Universe flat). Indeed in a two dimensional ($H_{0},w$) space the CMB constraint approximately lies on the region of constant large scattering surface distance as shown in \reffig{ellipse_w_H0}.
Notice that an external prior on $H_{0}$ two times more accurate than CMB S4 can improve the CMB constraint on $w$ by a factor of two.

@@

Much of the improvement that would come on neutrino masses and $w$ from $H_0$ or $\Omega_ch^2$ priors is anticipated to be provided by distance measurements made by DESI. But, additional $H_0$ measurements beyond DESI would help, as evidenced by the sharply falling curves in the left-most regions of \reffig{prior_omeganuh2-h}, \reffig{prior_omeganuh2-50-h} and \reffig{ellipse_w_H0}. Indeed, perhaps the most tantalizing finding is that future surveys beyond DESI and LSST \cite{lsst-dark-energy-science-collaboration:2012} that would tighten the constraints on, say, $n_s$ could indirectly lead to constraints in $N_{\rm eff}$ much less than a percent. Given the power of these constraints to probe physics beyond the Standard Model, such surveys seem particularly interesting to pursue.

>>>>

	
% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showpacs' option to make PACS codes appear
%  Add 'showkeys' option to make keywords appear
\documentclass[aps,prd,reprint,superscriptaddress,nofootinbib,floatfix]{revtex4-1}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{graphicx}
\usepackage{color}

\newcommand{\sd}[1]{\textcolor{red}{[{\bf SD}: #1]}}
\newcommand{\am}[1]{\textcolor{blue}{[{\bf AM}: #1]}}
\newcommand{\yp}[1]{\textcolor{green}{[{\bf YP}: #1]}}

\newcommand\refsec[1]{\S\ref{sec:#1}}
\newcommand\refeq[1]{Eq.~(\ref{eqn:#1})}
\newcommand{\refssec}[1]{section~\ref{subsec:#1}}
\newcommand{\reffig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\physrep}{Physics Reports}
\newcommand{\aapr}{A\&A Rev.}
\newcommand{\apjl}{ApJL}
\newcommand{\apjs}{ApJS}
\newcommand{\aap}{A\&A}
\newcommand{\mnras}{MNRAS}
\newcommand{\physrev}{Phys. Rev.}
\newcommand{\physrevlett}{Phys. Rev. Lett.}
\newcommand{\jcap}{{\em JCAP }}
\newcommand{\aj}{AJ }
\bibliographystyle{apsrev4-1}

@@

\section{Results \label{sec:results}}
In this section we present our main results. Starting from $\nu \Lambda$CDM we will then explore extensions to this model by considering constraints on $N_{\rm eff}$ and $w$.

@@

Small scale structure formation is suppressed if fast-moving neutrinos comprise a significant part of the matter budget. 
Indeed, below the free-streaming scale, the matter power spectrum is suppressed in the presence of
massive neutrinos by a factor $\Delta P/P\simeq -8f_{\nu}$ , where $f_{\nu} = \Omega_{\nu} / \Omega_{m}$ is the contribution of neutrinos to the total matter density.
This effect is probed by CMB lensing and, since the active neutrino number densities are known in the Standard Model, it can be directly transformed into a constraint on the sum of the neutrino masses $\sum m_\nu$. This is particularly exciting because we now know that neutrinos are massive, with a lower limit $\sum m_\nu>50$ meV that emerges from oscillation experiments (see \cite{bellini:2013} and references therein).

%==========
\begin{figure}[htbp]
\begin{center}
\includegraphics{{prior_omnuh2_lmin=4_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
\caption{Projected constraints on the sum of the neutrino masses (absolute on the left and relative on the right) from CMB-S4 and current BAO measurements as a function of priors on the other cosmological parameters. The {\it internal} (marginalized constraints from CMB-S4 and current BAO) constraints on each parameter are given as $\sigma_{\rm pipeline}$. The horizontal (blue) line correspond to CMB S4 experiment + Planck Pol + DESI.}
\label{fig:prior_omeganuh2-h}
\end{center}
\end{figure}
%==========
%==========

\begin{figure}[htbp]
\begin{center}
\includegraphics{{prior_omnuh2_notau_lmin=50_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
\caption{Same as Fig.~1, but with $l<50$ modes excluded from CMB-S4, instead coming from Planck. Note the importance of obtaining information on the optical depth $\tau$ if the low-$l$ modes are not measured accurately.}
\label{fig:prior-50}
\end{center}
\end{figure}
%==========

@@

\reffig{prior_omeganuh2-h} shows the projected errors on $\sum m_\nu$ from CMB-S4 (+Planck+current BAO) as a function of external priors on $H_{0}, \Omega_{c}h^{2}, n_s$ and $A_s$. 
Without any external priors (far right in figure), upcoming CMB experiments are poised to obtain 1-sigma limits on $\sum m_\nu$ of order 25 meV, corresponding to a 2-sigma detection even in the worst case scenario. Adding in external priors helps this significantly though, as discussed, for example, in \cite{di-valentino:2015,2013arXiv1309.5383A,pan:2015a,allison:2015}. 
There, the added prior was ``DESI BAO'', representing a measurement of distances as a function of redshift from the baryon acoustic oscillation feature probed by the planned experiment DESI \cite{levi:2013}. This is represented by the light blue band in our figures. In addition to showing the implications of introducing this fixed external constraint, \reffig{prior_omeganuh2-h} generalizes the extent to which external information will improve the neutrino mass constraints. For example, in \reffig{prior_omeganuh2-h} the red solid curve shows the improvement as additional information on the Hubble constant is applied. We can see, for example that, imposing {\it external} information at the level of $\sigma_{H_0}/H_0=0.5\%$ would improve the neutrino mass constraint to 17 meV. It is not surprising that this is close to the constraint obtained with DESI because the BAO distances in this model depends most sensitively on $\Omega_ch^2$ and $H_{0}$.
Similarly, adding to CMB S4 + BAO15 an external constraint on the dark matter density equal to its internal constraint of $0.4\%$ improves the projected neutrino mass constraint to 20 meV. These are multi-dimensional examples of the familiar fact that adding two sets of equal and indipendent constraints (in this case internal to S4+BAO15 and external) reduces the overall error by a factor of $\sqrt{2}$.

@@

Because of the presence of foregrounds, it may be difficult to get the very low-$\ell$ modes from the ground with CMB-S4, so in \reffig{prior-50}, we limit the CMB S4 modes to $\ell>\ell_{\rm min}=50$. For the large scale modes $\ell<\ell_{\rm min}=50$ we use Planck at the level of accuracy forecasted by the Planck Blue Book.  
Together with the importance of $H_{0}$ and $\Omega_ch^2$ previously observed, in this case external priors on $A_{s}$ and $\tau$ are found to be important. 
Indeed, as recently pointed out by \cite{allison:2015}, a degeneracy exists between $\sum m_\nu$ and $A_{s}$ and, as a consequence, between $\sum m_\nu$ and $\tau$. This can be understood as follows.
The CMB lensing reconstruction is noisy at scales larger than the neutrino free streaming scale (see Fig. 5 of \cite{2013arXiv1309.5383A}). For this reason it is hard to measure the unsuppressed CMB lensing power spectra that is then compared to the small scale suppressed one to constrain the values of $f_{\nu}$. The unsuppressed amplitude of the lensing spectrum is actually better constrained by the primordial amplitude $A_{s}$, which is probed by the primordial CMB. Unfortunately, the CMB spectra are sensitive to $A_{s}\exp^{-2\tau}$, so a measurement of the optical depth is required to infer $A_{s}$ and therefore to tighten the constraints on $\sum m_\nu$. If CMB S4 will not be designed to measure scales $\ell<50$ to avoid difficulties in dealing with atmospheric emission, an external prior on the optical depth will be crucial. This can be obtained from future 21 cm surveys \cite{liu:2015} or other CMB experiments, satellite like LiteBIRD \cite{matsumura:2014}, PIXIE\cite{kogut:2011}, and COrE/PRISM \cite{the-core-collaboration:2011,prism-collaboration:2013}, balloon-borne experiments like SPIDER \cite{filippini:2010}, EBEX \cite{reichborn-kjennerud:2010} or eventually ground experiments like CLASS \cite{essinger-hileman:2014}.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics{{prior_massless_neutrinos_lmin=4_lmax=4499_ndet=1000000_fsky=0.75}.pdf}
\caption{Constraints on $N_{\rm{eff}}$ (absolute on the left and and relative to its value on the right) as a function of priors on $\Omega _{b} h^{2}$, $\Omega _{c} h^{2}$ and $n_{s}$. The horizontal (blue) line correspond to CMB S4 experiment + Planck Pol + DESI.}
\label{fig:prior_neff}
\end{center}
\end{figure}

\reffig{prior_neff} shows projections for how well CMB-S4 will do at measuring $N_{\rm eff}$ as a function of priors on $H_{0}$, $\Omega_{c}h^{2}$ and $\Omega_{b}h^{2}$ (the sum of the neutrino masses must be included as a free parameter). With no priors on the other parameters, the projected $1$-sigma error is $0.016$, close to a 3-sigma detection of the expected deviation from $N_{\rm eff}=3$.

The sensitivity to relativistic degrees of freedom comes from the effect of extra species on the damping tail of the CMB anisotropies \cite{2013PhRvD..87h3008H}, both in temperature and polarization, so parameters that also strongly affect this part of the spectrum, like the slope $n_s$ and the baryon density $\Omega_bh^2$, are the most relevant to improve the $N_{\rm eff}$ constraint. As such, the blue and green dashed lines in \reffig{prior_neff} show that obtaining external priors on either of these would reduce the errors on $N_{\rm eff}$ to get close to a 4-sigma detection of the partial decoupling prediction. Besides futuristic high redshift 21 cm ideas, the best bet to improve the constraints on $n_s$ are probes that push to redshift $z>1$ such as the Lyman alpha forest. There are many more linear Fourier modes available at high $z$ that can be used to obtain the larger level arm from large to small scales that is essential to extract high precision measurements of $n_s$.

@@

\reffig{prior_w-h} shows that without any priors, CMB experiments (+BAO15) will do better than current constraints, which hover around 10\%. However, an external an external prior on $H_{0}$ two times more accurate than CMB S4 (i.e. $1\%$) can improve the CMB constraint on $w$ by a factor of two. 
This is primarily due to the fact that CMB constrains $w$ through a precise measurement of the comoving distance to the last scattering surface. However that can be kept fixed while varying $w$ by accordingly changing the value of $H_{0}$ (and $\Omega_{\Lambda}$ to keep the Universe flat). Indeed in a two dimensional ($H_{0},w$) space the CMB constraint approximately lies on the region of constant large scattering surface distance as shown in \reffig{ellipse_w_H0}.

@@

Much of the improvement that would come on neutrino masses and $w$ from $H_0$ or $\Omega_ch^2$ priors is anticipated to be provided by distance measurements made by DESI. But, additional $H_0$ measurements beyond DESI would help, as evidenced by the sharply falling curves in the left-most regions of \reffig{prior_omeganuh2-h}, \reffig{prior-50}, \reffig{prior_neff} and \reffig{prior_w-h}. Indeed, perhaps the most tantalizing finding is that future surveys beyond DESI and LSST \cite{lsst-dark-energy-science-collaboration:2012} that would tighten the constraints on, say, $n_s$ could indirectly lead to constraints in $N_{\rm eff}$ much less than a percent. Given the power of these constraints to probe physics beyond the Standard Model, such surveys seem particularly interesting to pursue.
